Debiased Calibration for Generalized Two-Phase Sampling

### 2025-01-20 final/paper/paper_ov.tex
Updated paper with Dr. Kim corrections.

\begin{abstract}
Generalized entropy calibration \cite{gneiting2007strictly} is a nonparametric
method to estimate survey weights without using the design weights in the
specified loss function. While \cite{kwon2024debiased} explore generalized
entropy calibration to estimate survey weights with known population totals, by
using a two-phase sampling framework, we extend their method to allow for
estimated population totals. We then extend this two-phase sampling framework to
two additional cases: first, when we have two non-nested samples, and
second, when we want to combine three or more samples. 
We apply our method to the Medicare Current Beneficiary Survey for which we
get more efficient estimates by combining this data set with population level
controls obtained from the Center of Medicare and Medicaid Services as well as
survey data from the U.S. Census' American Community Survey.
\end{abstract}
\newpage

@@

Combining information from multiple samples is an important practical problem
(\cite{yang2020statistical}, \cite{yang2023elastic}, \cite{dagdoug2023model}).
We want to incorporate information from external data sources to reduce the bias
in our estimates and improve the estimator's efficiency. For many problems, the
additional information consists of summary statistics with standard errors. The
goal of this project is to incorporate external information with existing data 
to create more efficient estimators using calibration weighting.

To model this scenario, we formulate the problem as a generalized two-phase
sample. The first phase sample consists of data from multiple sources, and the
second phase sample contains the existing data. We first consider the classical
two-pahse sampling setup where the second phase sample is a subset of the first
phase sample. Then we extend this setup to consider non-nested samples and
multiple non-nested samples.

\textcolor{red}{ Introduction is too short. We need to make 4-5 paragraphs. } 
\begin{enumerate}
\item Narrow down (like a funnel) to the subject that you want to address. 
\item Explain the context of the research (and cite a few key papers).
\item Explain what is proposed in the paper.  
\item Significance of the research (why this paper should be published)
\item Outline of the paper 
\end{enumerate}

@@

Consider a finite population of size $N$ 
with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the auxiliary variables and $y_i$ is the realization of the study variable of interest for element  $i$ in finite population. The auxiliary variables are cheap to measure but the measurement for the study variable is expensive. Thus,  two-phase sampling is popular as a tool for cost effective sampling technique. In the two-phase sampling, the first phase sample $A_1$  is selected from the target population to obtain the measurement of $\bx_i$ in the sample. The sampling design for the first-phase sample can be denoted by $P(A_1)$. After that, based on the observed auxiliary information in the first-phase  sample, the second-phase sample $A_1$ is selected from the first-phase sample to observe $y_i$. The sampling design for the second-phase sample is denoted as $P( A_2 \mid A_1)$.  
The
goal of two-phase sampling is to construct an estimator of $Y =  \sum_{i=1}^N y_i$ 
that uses both the observed information in the Phase 2 sample and also the extra
auxiliary information from $\bf X$ in the Phase 1 sample. How to combine  information efficiently is an important problem in two-phase sampling.

Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the second-phase sample. From the first-phase sample, we can compute $\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased estimator of $\mathbf{X}= \sum_{i=1}^N  \bx_i$. Also, from the second-phase sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \widehat{Y}_2 \right) = \sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as an unbiased estimator of $( \mathbf{X}, Y )$. Thus, we have three estimators for two parameters. The two-phase regression estimator can be  used to combine the information efficiently.  The two-phase regression estimator can be expressed as  
\begin{equation}
\hat Y_{\rm reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
\label{eq:1}
\end{equation} 
where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and  
$$
\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
 \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
$$ 
Regarding the choice of $q_i$, one can consider a linear regression model 
\begin{equation}
y_i = \bx_i^\top \bm \beta + e_i 
\end{equation}
where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known function. In this case,   $q_i=c_i$ is a reasonable choice.

One way to understand the regression estimator is to use the following decomposition: 
$$ 
\hat Y_{\rm reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
\text{Bias correction}}.
$$
The prediction term is obtained by finding the best linear unbiased predictor under the working superpopulation model. The bias correction term is used to make  the two-phase regression estimator be asymptotically design unbiased regardless of the underlying the superpopulation model.

%\textcolor{red}{I would move the calibration expression to the next subsection. }

Now, note that the two-phase regression estimator in (\ref{eq:1})  can be expressed as  $\hat Y_{\rm reg, tp} =
\sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where

$$
\hat w_{2i} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
\text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
\bf x_i / \pi_{1i}.
$$

This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than just squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ continuous.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others. The variance estimation is based on a linearization
that shows that minimizing the generalized loss function subject to the
calibration constraints is asymptotically equivalent to a regression estimator.

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. We would rather
have bias calibration separate from the minimizing the variance so that
we can control each in isolation. In
\cite{kwon2024debiased}, the authors show that for a generalized entropy
function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

@@

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method. Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is strictly convex, differentiable.  
We consider maximizing  
\begin{equation}\label{eq:primalloss}
  H(w) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
\end{equation} 
subject to the following constraints:

\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} \frac{\bf x_i w_{2i}q_i}{\pi_{1i}} = 
\sum_{i \in A_1} \frac{\bf x_iq_i}{\pi_{1i}}
\end{equation}

@@

The first constraint is the existing calibration constraint and the second
ensures that design consistency is achieved. 
The original method of \cite{kwon2024debiased} only considers known finite
population quantities on the right hand side of \eqref{eq:calconst1}.

@@

\begin{equation}\label{eq:primal}
  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
\end{equation}
where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
estimator of $Y_N$ is

@@

%\subsection{Theoretical Results}

We can understand the estimated weights, $\hat w_{2i}$ from 
Equation~\ref{eq:primal}
using the method of Legrange multipliers.
We need to minimize the Lagrangian function

\begin{equation}\label{eq:legragedc1}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
\end{equation}
where $\bm \lambda$ is a vector of Lagrange multipliers.
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero, yields the fact that $\hat w_{2i}$ satisfies

$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$

@@

The first result that we show is that the construction of the weights in
Equation~\ref{eq:primal} leads to a design consistent estimator 
$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.

\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
  Under some regularity conditions,

  $$\hat Y_{\rm DCE} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$

@@

  $$
  \hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{\rm DCE}(\hat{\bm \lambda}) + 
  \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i}(
  \hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi^*,
  $$
\textcolor{red}{(The above expression is not good. For example, the left-hand side is a function of $\bm \lambda$ and the right-hand side is a function of $\hat{\bm \lambda}$. Also, since we have $\bm \lambda^*=( \mathbf{0}, 1)$, the above expression can be simplified further. That is, there is no need to use $\bm \lambda^*$ in the above expression. ) }
  and

  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
  $$

   \begin{align*}
   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^T \bm \phi_1^* \\
   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
  \end{align*} 
\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
  and

  $$
  \begin{pmatrix}
  \bm \phi_1^* \\
  \phi_2^* 
  \end{pmatrix}
  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
  \begin{pmatrix}
  \bx_i \bx_i^T &   \bx_i g_i   \\
  g_i  \bx_i^T   & g_i^2     
  \end{pmatrix} \right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
  \begin{pmatrix}
    \bx_i \\ g_i 
  \end{pmatrix}y_i 
  $$

with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
\end{theorem}

\textcolor{red}{The presentation should be improved further. Since we

}

\begin{proof}
  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
  that it is asymptotically equivalent to a regression estimator. Using the
  method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
  minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
  The first order conditions show that

  $$
  \frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) w_{1i}q_i -
  w_{1i} \bm \lambda \bf z_i q_i = 0.
  $$

  Hence, $\hat w_{2i}(\bm \lambda) = g^{-1}(\bm \lambda^T \bf z_i)$ and
  $\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
  sample size gets large, we have $\hat w_{2i}(\hat{\bm \lambda}) \to
  d_{2i|1}$ which means that $\hat{\bm \lambda} \to \bm \lambda^*$ where $\bm
  \lambda^* = (\bf 0, 1)$. \textcolor{red}{(You should prove it rather than just saying it. ) }  
  Then using the linearization technique of
  \cite{randles1982asymptotic}, we can construct a regression estimator,

  $$
  \hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{\rm DCE}(\hat{\bm \lambda}) + 
  \left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
  \hat w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.
  $$

  Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \hat Y_{\rm DCE}(\hat{\bm
  \lambda})$ for all $\bm \phi$. We choose $\bm \phi^*$ such that

  \begin{align*}
    \bm \phi^*
    &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
    \right]^{-1}
    E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
    &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
    \right]^{-1}
    \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
  \end{align*}

  The final equality comes from the fact that 
  $E\left[\frac{\partial}{\partial \bm \lambda}
  \hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
  $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i}(\bm \lambda^*)$ is
  bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
  proves our result.
\end{proof}

\textcolor{red}{(Please move the proof to Appendix. Also, we need to discuss the implication of the theorem. Specifically, what is the expectation and variance of the linearization term? ) }

To construct asymptotically unbiased estimate of the variance we assume that the
sampling designs for $A_1$ and $A_2$ are measurable and we define the following
joint inclusion probabilities:

$\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.$

@@

\begin{proof}
  By Theorem~\ref{thm:dc1}, $\hat Y_{\rm DCE}$ is asymptotically equivalent to
  $\hat Y_{\ell, \rm{DCE}}$. Notice that we can define

>>>>

\begin{abstract}
Generalized entropy calibration \citep{gneiting2007strictly} is a nonparametric
method to estimate survey weights without using the design weights in the
specified loss function. \cite{kwon2024debiased} explore generalized
entropy calibration to estimate survey weights with known population totals.
By using a two-phase sampling framework, we extend this method to allow for
estimated population totals. We then generalize the two-phase sampling framework to
two additional cases: first, when we have two non-nested samples, and
second, when we want to combine three or more samples. 
We apply our method to the Medicare Current Beneficiary Survey for which we
get more efficient estimates by combining this data set with population level
controls obtained from the Center of Medicare and Medicaid Services as well as
survey data from the U.S. Census' American Community Survey.
\end{abstract}

@@

Government surveys in the United States provide the public with diverse
information about a range of population characteristics. These include health
indicators from the National Health and Nutrition Examination Survey, population
and housing data from the American Community Survey, and labor force statistics
from the Current Population Survey. All of these surveys are probability
surveys. Probability sampling is considered the
gold-standard for survey sampling. However, obtaining a large probability
sample is getting more difficult because people are more likely to choose not to
respond. One solution to this problem is data integration. Instead of analyzing
a probability sample by itself, we can incorporate information from other
probability samples in order to have an analysis with higher statistical power.

The goal of statistical data integration is to combine information from one
survey with the variable of interest to another survey with
shared covariates. 
This integration usually involves one of two contexts: incorporating item
level responses or using summary statistics. These mirror the micro and macro
approaches described in \cite{yang2020statistical} respectively. The main
technique to incorporate item level responses is mass imputation. With access to
individual item level responses, the traditional approach is to build a
predictive model of the variable of interest from the covariates shared between
the two surveys and predict the missing outcome variable in the survey without
the variable of interest. An early use of mass imputation is found in
\cite{breidt1996two} where they use mass imputation to improve estimation in
the National Resources Inventory survey. Later, \cite{kim2012combining} use a
parametric model to develop a novel resampling technique to construct confidence
intervals of population level characteristics. \cite{chen2022nonparametric} use
a nonparametric model to estimate population characteristics. In the macro
approach, one uses summary statistics from other studies to adjust the current
estimation. This is commonly done using generalized least squares (GLS).
\cite{hidiroglou2001double} apply the GLS technique to non-nested two-phase
samples. \cite{merkouris2004combining} use GLS with multiple surveys and showed
an optimality result. This was later extended to area level domains in
\cite{merkouris2010combining}.

In this paper, we incorporate summary statistics of other surveys by
using a generalized two-phase sample similar to the GLS method. The first phase
sample consists of data
from multiple sources, and the second phase sample contains the existing
data. Unlike previous data integration methods, however, we also include a
pre-chosen function of the second phase design weights like
\cite{kwon2024debiased}. This allows us to develop theory showing that our
estimates are asymptotically design unbiased and that we have valid confidence
intervals. Moreover, we can show that this technique outperforms traditional
regression methods in finite samples.

\textcolor{red}{(I am not sure whether the current introduction provides a good guide for the main contents of the paper. The introduction does not match  well with the main contents of the paper.)
}

This paper is organized as follows. Section 2 describes the basic setup of data
integration using summary statistics. \textcolor{red}{(This is not true.)} Section 3 details the generalized
two-phase sampling framework and provides theory for our estimator to be design
unbiased and the variance estimation to be valid. In Section 4, we extend this
framework first to non-nested two-phase sampling and second to multiple
samples. Section 5 provides several simulation studies which show that our
method works, and Section 6 applies the method to the Medicare Current
Beneficiary Survey.

@@

Consider a finite population, $U = \{1, \dots, N\}$, of size $N$ 
with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the
auxiliary variables and $y_i$ is the realization of the study variable of
interest for element  $i$ in the finite population. The auxiliary variables are
cheap to measure but the measurement for the study variable is expensive.
Two-phase sampling is a popular method for cost effective sampling.
In two-phase sampling, the first phase sample $A_1$  is selected from the
target population to obtain the measurement of $\bx_i$ in the sample. The
sampling design for the first-phase sample can be denoted by $P(A_1)$. After
that, based on the observed auxiliary information in the first-phase  sample,
the second-phase sample $A_2 \subseteq A_1$ is selected from the first-phase
sample to observe $y_i$. The sampling design for the second-phase sample is
denoted as $P( A_2 \mid A_1)$.  
The
goal of two-phase sampling is to construct an estimator of $Y_N =  \sum_{i=1}^N y_i$ 
that uses both the observed information in the Phase 2 sample and also the 
auxiliary information from $\bf X \in \R^p$ in the Phase 1 sample. How to
combine information efficiently is an important problem in two-phase sampling.

Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
second-phase sample. From the first-phase sample, we can compute
$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased
estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
\sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as
an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
We therefore have three estimators for
two parameters. The two-phase regression estimator can be  used to combine this
information efficiently and it can be expressed as  
\begin{equation} \label{eq:1}
\hat Y_{\rm reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
\end{equation}  
where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and

$$
\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
 \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
$$

Regarding the choice of $q_i$, one can consider a linear regression model

\begin{equation}
y_i = \bx_i^\top \bm \beta + e_i 
\label{regression}
\end{equation}

\textcolor{red}{(Why do you have a space here?) }
where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
function. In this case,  $q_i = c_i$ is a reasonable choice. 
One way to understand the two-phase regression estimator is to use the following
decomposition:

$$ 
\hat Y_{\rm reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{\bf x_i^\top \hat{\bm \beta_q}}{\pi_{1i}}}_{
  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
\text{Bias correction}}.
$$
The prediction term is obtained by finding the best linear unbiased predictor
under the working superpopulation model. The bias correction term 
makes the two-phase regression estimator asymptotically design unbiased
regardless of the underlying the superpopulation model. 
Note that the two-phase regression estimator in Equation~\ref{eq:1} can be
expressed as  $\hat Y_{\rm reg, tp} =
\sum_{i \in A_2} \pi_{1i}^{-1} \hat w_{2i} y_i$ where

$$
\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
\text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
\bf x_i / \pi_{1i}.
$$

This means that $\hat Y_{\rm reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ being continuous.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others \citep{schennach2007point}.

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. We would rather
have bias calibration separate from the minimizing the variance so that
we can control each in isolation. 
\cite{kwon2024debiased} shows that for a generalized entropy
function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

@@

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
strictly convex and differentiable. We consider maximizing

\begin{equation}\label{eq:primalloss}
  H(w_2) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
\end{equation}

subject to the following constraints:
\textcolor{red}{(Please do not put a space between a displayed equation and the following sentence, unless you want to make it as a separate paragraph. ) }

\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} \frac{\bf x_i w_{2i}}{\pi_{1i}} = 
\sum_{i \in A_1} \frac{\bf x_i}{\pi_{1i}}
\end{equation}

@@

The first constraint is the usual calibration constraint motivated from the
regression model in (\ref{regression}) and the second
ensures that design consistency is achieved. 
The original method of \cite{kwon2024debiased} only considers known finite
population quantities on the right hand side of \eqref{eq:calconst1}.

@@

\begin{equation}\label{eq:primal}
  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
\end{equation}

where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
estimator of $Y_N$ is

@@

\subsection{Theoretical Results}

We can understand the estimated weights, $\hat w_{2i}$ from
Equation~\ref{eq:primal} using the method of Lagrange multipliers. We need to
minimize the Lagrangian function

\begin{equation}\label{eq:legragedc1}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  + \bm \lambda^\top \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
\end{equation}

where $\bm \lambda$ is a vector of Lagrange multipliers.
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero, we can obtain

$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$

where $g(w) = \partial G / \partial w$ and 
$\hat{\bm \lambda}$ is the solution to

@@

The following theorem shows that the construction of the weights in
Equation~\ref{eq:primal} leads to a design consistent estimator 
$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.

\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
  Under the regularity conditions stated in the appendix

  $$
  \hat Y_{\rm DCE} = \hat Y_\ell(\bm{\lambda}^*, \bm \phi^*) + O_p(N / n_2)
  $$

@@

  $$
  \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\bm \phi^*,
  $$

  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
  $$

%   \begin{align*}
%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
%   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
%  \end{align*} 
%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
%  and  
%
%  $$
%  \begin{pmatrix}
%  \bm \phi_1^* \\
%  \phi_2^* 
%  \end{pmatrix}
%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
%  \begin{pmatrix}
%  \bx_i \bx_i^\top &   \bx_i g_i   \\
%  g_i  \bx_i^\top   & g_i^2     
%  \end{pmatrix} \right]^{-1}
%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
%  \begin{pmatrix}
%    \bx_i \\ g_i 
%  \end{pmatrix}y_i 
%  $$

% with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
\end{theorem}

See the appendix for the proof of Theorem~\ref{thm:dc1}. Notice that this
theorem implies
that $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
the same asymptotic expectation and variance, 
which means that we can estimate
the variance of $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ to get an
estimate of the variance of $\hat Y_{\rm DCE}$. The basic idea of 
Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
method is equivalent to a regression estimator. By constraints 
(\ref{eq:calconst1}) and (\ref{eq:calconst2}), 
$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent, and the
variance is asymptotically equivalent to the two-phase regression estimator. 
We can estimate $\bm \phi^*$ with
\begin{equation}
\hat{\bm \phi} = 
\left[\sum_{i \in A_2} 
w_{1i}\frac{\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
\sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
\label{eq:8}
\end{equation}
Therefore, the proposed calibration estimator in (\ref{eq:primal}) is asymptotically equivalent to the two-phase regression estimator given by 
  $$
  \hat Y_{\rm tp, reg}  = 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\hat{\bm \phi},
  $$
  where $\hat{\bm \phi}$ is defined in (\ref{eq:8}).

For specific sample designs, we can construct the optimal entropy function
$G(\cdot)$. From Theorem~\ref{thm:dc1}, the choice of entropy function affects
the variance of the debiased calibration estimator through the form of $g'(d_{2i|1})$
in $\bm \phi^*$. Consider a class of generalized regression estimators,

$$
\mathbf{B} = 
\left\{\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \gamma + 
\sum_{i \in A_2} \frac{w_{1i}}{\pi_{2i|1}}(y_i - \bf{z}_i^\top \bm \gamma):
\bm \gamma \in \R^{p+1}\right\}
$$

Then the design optimal estimator is

$$
\hat \theta_{\ell, \rm{opt}} = 
\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
\frac{w_{1i}}{\pi_{2i|1}}\sum_{i \in A_2} (y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
$$

$$
\bm \phi_{\rm opt} = 
\Var\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i\right)^{-1} 
\Cov\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i, \sum_{i \in A_2}, d_{2i|1} y_i \right).
$$

We assume that both $A_1$ and $A_2$ are Poisson samples. Then,

$$
\bm \phi_{\rm opt} = 
\left\{
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i}\bf{z_i}^\top
\right\}^{-1}
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i} y_i.
$$

To find the optimal entropy function, we solve the differential equation

$$
\frac{1}{d_{2i|1} g'(d_{2i|1})} = d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1).
$$

Using the fact that $(y(xy-1))^{-1} = -y^{-1} + x(xy-1)^{-1}$, the optimal
entropy function is

$$
G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}w_{1i} - 1).
$$

To construct an asymptotically unbiased estimate of the variance we assume that
the sampling designs for $A_1$ and $A_2$ are measurable and we define the
following joint inclusion probabilities:

$$
\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.
$$

The next theorem discusses how to compute an estimate for the variance of 
$\hat Y_{\rm DCE}$. For the proof see the appendix.

@@

\section{Extensions}

\subsection{Non-nested Two-Phase Sampling}
